---
title: "Video_Project"
output: html_document
author: "Boyu Zheng"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
##Load Data

```{r}
data <- read.csv("Video_project_dataset.csv")
summary(data)
```


##Simple EDA

Except FEMALE (female proportion) and PCT_BA (percentage of people with bachelor degree in student's neighbourhood), most of variables are skewed. Extremely the demographic and income variables like PCT_BLACK, PCT_ASIAN, PCT_ASIAN, POVERTY_RATE.

Demographic variables also have many outliers. COSTT4_A (Cost of Attendance) and PCT_BA have very few outliers.

```{r}
par(mfrow=c(3,3))
hist(data$ADM_RATE, main="Histogram of Admission Rate", ylab="Admission Rate", breaks=20)
hist(data$COSTT4_A, main="Histogram of Average Cost of Attendance", ylab="Average Cost of Attendance",breaks=20)
hist(data$AVGFACSAL, main="Histogram of Average Faculty Salary", ylab="Average Salary",breaks=20)
hist(data$FEMALE, main="Histogram of Female Proportion", ylab="Percentage of Female Student",breaks=20)
hist(data$PCT_BA, main="Histogram of Education Level", ylab="Percentage of people with bachelor degree, over 25",breaks=20)
hist(data$PCT_BORN_US, main="Histogram of Porportion of people born in US", ylab="Percentage of people born in US",breaks=20)
hist(data$PCT_BLACK, main="Histogram of Porportion of Black", ylab="Percentage of Black people in Student's Neighbourhood",breaks=20)
hist(data$PCT_ASIAN, main="Histogram of Porportion of Asian", ylab="Percentage of Asian people in Student's Neighbourhood",breaks=20)
hist(data$POVERTY_RATE, main="Histogram of Poverty Rate", ylab="Poverty Rate",breaks=20)
```

```{r}
par(mfrow=c(3,3))
boxplot(data$ADM_RATE, main="Histogram of Admission Rate", ylab="Admission Rate", horizontal=TRUE)
boxplot(data$COSTT4_A, main="Histogram of Average Cost of Attendance", ylab="Average Cost of Attendance",horizontal=TRUE)
boxplot(data$AVGFACSAL, main="Histogram of Average Faculty Salary", ylab="Average Salary",horizontal=TRUE)
boxplot(data$FEMALE, main="Histogram of Female Proportion", ylab="Percentage of Female Student",horizontal=TRUE)
boxplot(data$PCT_BA, main="Histogram of Education Level", ylab="Percentage of people with bachelor degree, over 25",horizontal=TRUE)
boxplot(data$PCT_BORN_US, main="Histogram of Porportion of people born in US", ylab="Percentage of people born in US",horizontal=TRUE)
boxplot(data$PCT_BLACK, main="Histogram of Porportion of Black", ylab="Percentage of Black people in Student's Neighbourhood",horizontal=TRUE)
boxplot(data$PCT_ASIAN, main="Histogram of Porportion of Asian", ylab="Percentage of Asian people in Student's Neighbourhood",horizontal=TRUE)
boxplot(data$POVERTY_RATE, main="Histogram of Poverty Rate", ylab="Poverty Rate",horizontal=TRUE)
```

##Simple Linear Regression

Except Cost of Attendance (COSTT4_A) and Education level (PCT_BA), we see no significantly visible linear relationship between the variable and admission rate. Both COSTT4_A and PCT_BA have negative relationship with admission rate. 

```{r}
plot(x=data$COSTT4_A,y=data$ADM_RATE,xlab="Average Cost of Attendance", ylab="Admission Rate", main="Scatter plot for Cost vs Admission Rate")
plot(x=data$PCT_BA,y=data$ADM_RATE,xlab="Percentage of People with Bachelor Degree", ylab="Admission Rate", main="Scatter plot for Education Level vs Admission Rate")
plot(x=data$PCT_BORN_US,y=data$ADM_RATE,xlab="Percentage of People Born in US", ylab="Admission Rate", main="Scatter plot for US Proportion vs Admission Rate")
plot(x=data$POVERTY_RATE,y=data$ADM_RATE,xlab="Poverty Rate", ylab="Admission Rate", main="Scatter plot for Poverty Rate vs Admission Rate")
plot(x=data$PCT_BLACK,y=data$ADM_RATE,xlab="Percentage of Black People", ylab="Admission Rate", main="Scatter plot for Black vs Admission Rate")
plot(x=data$PCT_ASIAN,y=data$ADM_RATE,xlab="Percentage of Asian People", ylab="Admission Rate", main="Scatter plot for Asian vs Admission Rate")
```

## One multiple linear regression model, with F test

Result shows significant F test result, which means linear relationship exists. However, R-squared is small and some variables (like REGION and PCT_ASIAN) are not significant. We may need a partial F test.

```{r}
mod1 <- lm(ADM_RATE ~ REGION+PBI+WOMENONLY+COSTT4_A+PCT_BA+PCT_BORN_US+PCT_BLACK+PCT_ASIAN+POVERTY_RATE, data=data)
summary(mod1)
```

